---
title: "BactaxR"
output: html_notebook
---

## Install a modified version of `bactaxR`

Follow the installation instructions in the original repository, https://github.com/lmc297/bactaxR. Then, after installing the dependencies, install the library from my repository:

```{r}
# library(devtools)
# install_github("matrs/bactaxR")
```


This version adds 1 line to the original code, to make it possible to run it with fastani tables which have missing pairs (`fastani` doesn't report pairs when the ANI is below a threshold, which is hard-coded)

```{r}
library(bactaxR)
```

## Read the output from `fastani`

Create an `ani` object.
```{r}
ani <- read.ANI(file = "../g_akk_gtdb.tsv")
h <- ANI.histogram(bactaxRObject = ani, bindwidth = 0.1)
h
```
## Generate multiple clusters' tables using different ANIs thresholds

using multiple ANI thresholds, we can create multiple clusterings. See what the `dend` object has:

```{r}
dend <- ANI.dendrogram(bactaxRObject = ani, ANI_threshold = 95, label_size = 0.5)
head(dend$cluster_assignments)
```

Create one table per ANI:
```{r eval=FALSE}
# from ANI 95 to 97 in 0.5 steps
for (i in seq(95, 97, 0.5)){
  cat(i,'\n')
  dend_g_akk <- ANI.dendrogram(bactaxRObject = ani, ANI_threshold = i, label_size = 0.1)
  write.table(dend_g_akk$cluster_assignment, file=paste0('g_akk_clusters_', i, '.tsv'), 
            row.names = FALSE, quote=FALSE, sep='\t')
}
```

These tables can be used to parse multiple species trees and see if a particular ANI coincides the best with the tree. Also, we can generate cluster visualizations:

```{r}
metadata <- dend_g_akk$cluster_assignments$Cluster
names(metadata) <- dend_g_akk$cluster_assignments$Genome
g_akk_ani_clust_plot <- ANI.graph(bactaxRObject = ani, ANI_threshold = 95,
          metadata = metadata,
          legend_pos_x = -1.5, show_legend = T, graphout_niter = 1000000, 
          legend_ncol = 1, edge_color = hcl(h = 168, c = 5, l = 58, alpha = 0.1), legend_cex = 1.2)
```
